---
title: "IS 607 Final Project"
author: "Puneet Auluck"
date: "December 13, 2015"
output: html_document
---

Objective
========================================================
My objective for this project is to find best deals on NY hotels in January 2016 on popular travel sites.  For now, I have decided to use the first month of the year but it can be easily tweaked to view other months. 

### Load packages
```{r message=F, warning=F}
# Load packages needed
library(RCurl)
library(XML)
library(stringr)
library(scatterplot3d) 
library(plotrix) 
library(ggplot2)
library(dplyr)
```

### Create getHotelBookings() wrapper function
```{r}
# getHotelBookings wrapper function to get bookings
getHotelBookings <- function (url, date){
  #print(url)
  results <- getURL(url)
  results_tree <- htmlParse(results)
  hotel_db_tmp <- as.character()
  hotels <- as.character()
  
  hotels <- xpathSApply(results_tree,"//div[@class='item_prices cf']//h3",xmlValue)
  
  # Hotel option deal 1
  site_1 <- xpathSApply(results_tree,"//div[@class='item_main']//li[1]//em",xmlValue)
  price_1 <- xpathSApply(results_tree,"//div[@class='item_main']//li[1]//strong",xmlValue)
  hotel_db_tmp <- cbind(hotels,site_1,price_1,date)
  
  # Hotel option deal 2
  site_2 <- xpathSApply(results_tree,"//div[@class='item_main']//li[2]//em",xmlValue)
  price_2 <- xpathSApply(results_tree,"//div[@class='item_main']//li[2]//strong",xmlValue)
  hotel_db_tmp <- rbind(hotel_db_tmp, cbind(hotels,site_2,price_2,date))
  
  # Hotel option deal 1
  site_3 <- xpathSApply(results_tree,"//div[@class='item_main']//li[3]//em",xmlValue)
  price_3 <- xpathSApply(results_tree,"//div[@class='item_main']//li[3]//strong",xmlValue)
  hotel_db_tmp <- as.data.frame(rbind(hotel_db_tmp, cbind(hotels,site_3,price_3,date)))
  
  return(hotel_db_tmp)
}
```

### Web scape bookings for each day in Jan '16. The pages scraped are saved to csv file.  The csv file is then loaded to github for further analysis
```{r}
# Construct part of URL
base_url <- "http://www.trivago.com/?iPathId=34812&iGeoDistanceItem=0&aDateRange%5Barr%5D=2016-"
date_end <- "&aDateRange%5Bdep%5D=2016-"
end_url <- "&iRoomType=7&cpt=3475703%2C199&aHotelTestClassifier=&aPriceRange%5Bfrom%5D=0&aPriceRange%5Bto%5D=0&iIncludeAll=0&aPartner=&iViewType=0&isRetina=false&bIsSeoPage=false&bIsSitemap=false&"

# Loop through each day of bookings in January 2016
dates <- 1:30
hotel_db <- as.character()
for (i in dates){
  url_tmp <- paste(base_url,"01-",i,date_end,"01-",i+1, end_url,sep="")
  hotel_date_added <- getHotelBookings(url_tmp,paste("2016-01-",formatC(i, width=2, flag="0"),sep=""))
  hotel_db <- as.data.frame(rbind(hotel_db,hotel_date_added))
}

# Don't forget 01/31-02/01 data
url_tmp <- paste(base_url,"01-31",date_end,"02-01",end_url,sep="")
hotel_date_added <- getHotelBookings(url_tmp,"2016-01-31")
hotel_db <- as.data.frame(rbind(hotel_db,hotel_date_added))

colnames(hotel_db) <- c("hotel","site","price","check_in_date")
head(hotel_db)
```


### Clean up
```{r}
hotel_db_cleaned <- hotel_db
hotel_db_cleaned$hotel <- str_trim(hotel_db_cleaned$hotel)
hotel_db_cleaned$site <- str_trim(hotel_db_cleaned$site)
hotel_db_cleaned$price <- gsub(".*\\$","",hotel_db_cleaned$price)
hotels <- hotel_db_cleaned
head(hotels)
```

# Write to csv
```{r, eval=FALSE}
# Once the pages are scraped, write to csv file that can be used for analysis
# It will be placed on github to be read from there
write.csv(hotels, file = "c:\\607_final_project\\hotel_data.csv", row.names = FALSE)
```

### For rmd knitting process to be faster, I stored csv file to github and used the link from there.  
```{r message=F, warning=F}
# Read csv file from github
# Comment out if pages need to be scraped again
#csvfile <- "https://raw.githubusercontent.com/pauluck/R-Workshop/master/hotel_data.csv"
#hotels <- read.csv(csvfile, header = TRUE, sep = ",", na.strings=c("NA", "NULL"))
```

```{r message=F, warning=F}
# Filter out listings to subset of priced ranging from $10-$200
hotels$price <- as.numeric(as.character(hotels$price))
hotels$site <- as.character(hotels$site)
hotels$check_in_date <- as.character(hotels$check_in_date)
hotels$hotel <- as.character(hotels$hotel)
hotels_1 <- hotels[hotels$price>10 & hotels$price<200,]
head(hotels_1)
```

### Find the average of booking rate in January 
```{r message=F, warning=F}
# Find average price of the month - do not count null or n/a values
avg_price <- hotels_1 %>% select(price) %>% summarise(avg=mean(price, na.rm=TRUE))
avg_price
```

### Plot prices for each day scraped
```{r message=F, warning=F}
# plot all hotel prices for each day
hotel_date_split <- hotels %>% select(site,price,hotel,check_in_date) %>% mutate(date = gsub("(2016-01-)(\\d\\d)","\\2",check_in_date))
p1 <- ggplot(data = hotel_date_split, aes(x = hotel, y = price, group=hotel))
p1 + geom_point(outlier.shape = 3) + theme(axis.ticks = element_blank(), axis.text.x = element_blank()) + facet_wrap(~date)
```

The price appears to be steady for each day.  

### Find the best deals for each day
```{r message=F, warning=F}

# Find minimum price of most popular hotels for each day
best_deal <- as.data.frame(hotels_1 %>% filter(price>10) %>% group_by(check_in_date) %>% 
                    summarize(price=price[which.min(price)],
                              hotel=hotel[which.min(price)],
                              site=site[which.min(price)]))
head(best_deal)
```

### Which hotel has the best deals in January?
```{r message=F, warning=F}
# Which hotel has the best deals in January
m <- ggplot(data=best_deal, aes(best_deal$hotel))                
m + geom_histogram(aes(fill = ..count..), binwidth = 0.1) +
  ggtitle("Hotels with Best Deals") +
  labs(x="Hotels",y="Frequency") + 
  theme(plot.title = element_text(family = "Trebuchet MS", color="#666666", face="bold", size=16, hjust=0)) +
  theme(axis.title = element_text(family = "Trebuchet MS", color="#666666", face="bold", size=16)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
 

### Which site delivers the most best deals?
```{r message=F, warning=F}
# Which site delivers the most best deals
best_site_feq <- best_deal %>% select(site,price) %>%
  group_by(site) %>%
  summarise(count=n()) 
best_site_pct <- best_site_feq %>% mutate(pct=paste(site," - ", formatC(count/sum(count)*100,digits=2),"%", sep=""))

pie3D(best_site_pct$count,labels=best_site_pct$pct,explode=0.1,
  	main="Best Hotel Sites")
```


```{r message=F, warning=F}
# Plot all best deals
dev.off
with(best_deal, {
  s3d <- scatterplot3d(as.factor(hotel), as.factor(check_in_date), price,        
                       color="green", pch=19,        
                       type="h",                    
                       main="Best Deals Price for Each Day",
                       xlab="hotels",
                       ylab="Dates",
                       zlab="Price")
  s3d.coords <- s3d$xyz.convert(site, check_in_date, price) 
  text(s3d.coords$x, s3d.coords$y, labels="")
})
```




