---
title: "607 Project 1"
author: "Puneet Auluck"
date: "September 24, 2015"
output: html_document
---
***
<span style="color:blue">**Load the packages for later use**</span><br>

```{r message=F, warning=F}
library(dplyr) # Selecting and filtering
library(magrittr) # Pipes
library(knitr) # kable
library(stringr) # String manipulation
```

<span style="color:blue">**Read file as a string**</span><br>
```{r message=F, warning=F}
RawFile <- paste(readLines("tournamentinfo.txt"), collapse = "\n")

# This is a partial view of how the string looks like
paste(readLines("tournamentinfo.txt", n=20), collapse = "\n")

```

<span style="color:blue">**Clean up and create a table for each player per row**</span><br>
```{r}
# remove unnecessary characters and merge 2 lines of player stats into one
splitRawFile <- gsub("\n\\s+|--+","",RawFile) %>% # remove lines containing #----------#
                gsub("\\|\\s*\n", "\n", .) %>% # merge each player stats into 1 line
                strsplit(.,"\n") # split the file containing one player per line

# read into a table
df_chess <- read.table(textConnection(splitRawFile[[1]]), sep="|", skip = 1)
kable(head(df_chess))
```

<span style="color:blue">**Get subset of columns that are needed**</span><br>
```{r}
# Only need first 12 columns of data and give them names
cnames <- c("PlayerNum","Player","State","TotalPoints", "R1","R2","R3","R4","R5","R6","R7","Rating")
df_chess_subset <- df_chess[c(1,2,11,3:10,12)]
colnames(df_chess_subset) <- cnames
kable(head(df_chess_subset))
```

<span style="color:blue">**Trim leading and trailing whitespace in Player names and State**</span><br>
```{r}
df_chess_subset$State <- str_trim(df_chess_subset$State)
df_chess_subset$Player <- str_trim(df_chess_subset$Player)

```

<span style="color:blue">**Clean up Rating column**</span><br>
```{r}
df_chess_subset$Rating <- as.character(gsub("^.*R:\\s+(\\d+)(\\D|\\s).*$","\\1",df_chess_subset$Rating))
kable(head(df_chess_subset))
```

<span style="color:blue">**Create another data frame containing only the Player number and Rating**</span><br>
```{r}
df_chess_rating <- df_chess_subset[c("PlayerNum","Rating")]
head(df_chess_rating)

```


<span style="color:blue">**Replace Player number in R1:R7 to their Ratings, respectively**</span><br>
```{r}
df_chess_subset[5:11] <- lapply(df_chess_subset[5:11], function(x) 
                                df_chess_rating[as.numeric(str_extract_all(x,"\\d{1,}")),"Rating"])
kable(head(df_chess_subset))
```

<span style="color:blue">**Now calculate the average and extract columns that are needed**</span><br>
```{r}
df_chess_avg <- df_chess_subset %>% 
                group_by(Player) %>% 
                mutate(Avg = round(mean(as.numeric(c(R1,R2,R3,R4,R5,R6,R7)), na.rm=T))) %>%
                select(Player, State, TotalPoints, as.numeric(Rating), Avg)

kable(head(df_chess_avg))
```

<span style="color:blue">**Write data frame to .csv file and read it from there as well**</span><br>
```{r}
# Write .csv
write.csv(df_chess_avg, file = "chess.csv", row.names = FALSE, quote = FALSE)

# Read .csv
kable(head(read.csv(file = "chess.csv", row.names = 1)))
```


***