---
title: "IS 607 Project 2"
author: "Puneet Auluck"
date: "October 6, 2015"
output: 
  html_document:
    toc: true
---


```{r message=F, warning=F}
# Get the packages we need.
require(tidyr)
require(dplyr)
require(knitr)
require(stringr)
require(ggplot2)
```


## <span style="color:#B05F3C;">**Data Set 1 - School Grades - By Puneet Auluck (Myself)**</span>
<div style="border: 3px solid #B05F3C;">
```{r}
# Read csv file
grades <- read.csv(file = "C:\\project2\\grades.csv", stringsAsFactors = FALSE)
kable(head(grades))
```

</div>

### <span style="color:#B05F3C">1. Is there a correlation between how well a student did in the class and their school attendance?</span>

<div style="border: 3px solid #B05F3C;">

<span style="color:green">&nbsp;**The original grades.csv wide format is made into long by gathering Subject and corresponding Grade.**</span>
```{r}
grades_l <- grades %>% gather("Subject", "Grades", 2:6)
kable(head(grades_l %>% arrange(desc(Days.Present))))
```

</div> <br>
<div style="border: 3px solid #B05F3C;">

<span style="color:green">&nbsp; **Let's find grade average for each student.**</span>



```{r}
grades_avg <- grades_l %>% group_by(Student,Days.Present) %>%
              summarise(Avg=mean(Grades))
              
kable(grades_avg %>% arrange(desc(Days.Present)))

```

<span style="color:green">&nbsp;**Student, Winston Cadiz, with highest attendance has grade average of 74.6 and student, Ronald Rosario, with lowest attendance has 78.4.  Based on grades.csv table, there no indication of a positive correlation between the days a student is present in school and how well he/she does academically.**</span>

<span style="color:green">&nbsp;**Here is a graph to visually represent the conclusion.**</span>


```{r  fig.height=4}
grades_plot <- ggplot(grades_avg, aes(y=Avg, x=Days.Present)) + geom_point(shape=1)
grades_plot + geom_smooth()

```

</div>

### <span style="color:#B05F3C">2. How are the grades by subject?<span>
<div style="border: 3px solid #B05F3C;">

```{r  fig.height=4}
grades_plot2 <- ggplot(data= grades_l, aes(y=Grades, x=Days.Present, fill=factor(Subject))) + geom_point() 
grades_plot2 + geom_smooth() + facet_wrap(~Subject)

```

<span style="color:green">&nbsp; **It appears, aside from Science and Technology, the grades are affected by the number of days a student is present in class.**</span>
</div>

## <span style="color:#9191FF;">**Data Set 2 - Living Former Presidents - By Youqing Xiang**</span>

<div style="border: 3px solid #9191FF;">

```{r}
# Read csv file
presidents <- read.csv(file = "C:\\project2\\presidents.csv", stringsAsFactors = FALSE)
kable(presidents)
```

</div> <br>
<div style="border: 3px solid #9191FF;">
<span style="color:green">&nbsp;**Tidy data**</span>

```{r}
presidents_tidy <- presidents %>%
                   separate(OfficeTerm, into=c("OfficeTermStart", "OfficeTermEnd"), sep="-") %>%
                   mutate(OfficeYears = as.numeric(OfficeTermEnd)-as.numeric(OfficeTermStart)) %>%
                   separate(BirthDate, into=c("BirthDate", "Age"), sep="\\(age ")
presidents_tidy$Age <- str_replace(presidents_tidy$Age, "\\)$","")
kable(presidents_tidy)
```

</div>

### <span style="color:#9191FF"> 1. Give the list of presidents who had one term(4 years in the office) and presidents who had two terms(8 years in the office)</span>

<div style="border: 3px solid #9191FF;">

```{r}
kable(presidents_tidy[presidents_tidy$OfficeYears == 4,])
kable(presidents_tidy[presidents_tidy$OfficeYears == 8,])
```

</div>

### <span style="color:#9191FF">2. Which president is the oldest one?</span>

<div style="border: 3px solid #9191FF;">

```{r}
kable(presidents_tidy %>% select(President,Age) %>% filter(Age == max(presidents_tidy$Age)))

```

</div>

### <span style="color:#9191FF">3. What is the average age of all the presidents?</span>
<div style="border: 3px solid #9191FF;">

```{r}
mean(as.numeric(as.character(presidents_tidy$Age)))
```

</div>

## <span style="color:#4A9586;">**Data Set: 3 - U.S. International Trade - By Jason Joseph**</span>
<div style="border: 3px solid #4A9586;">

```{r}
# Read csv file
trade <- read.csv(file = "C:\\project2\\trades.csv", stringsAsFactors = FALSE)
colnames(trade) <-c("Period","Type","2013-Exports","2013-Imports","2014-Exports","2014-Imports","2015-Exports","2015-Imports")
kable(trade)
```

</div> <br>
<div style="border: 3px solid #4A9586;">
<span style="color:green">&nbsp;**Drop unwanted rows and add missing data in first column.**</span>

```{r}

trade[c(seq(4, 13, 3)), 1] <- trade[c(seq(3,12,3)), 1] 
trade <- trade[c( -1, -2,c(seq(-11, -5, 3))),] #drop unwanted rows
kable(trade)
```

</div><br>

<div style="border: 3px solid #4A9586;">
<span style="color:green">&nbsp;**Tidy data**</span>
```{r}
trade_tidy <- trade %>% gather("TradeType", "Trades", 3:8) %>%
              separate(TradeType, into=c("Year","TradeType"), sep="-")

trade_tidy$Trades <- as.numeric(gsub(",", "", as.character(trade_tidy$Trades)))
trade_tidy$Type <- str_trim(trade_tidy$Type)
kable(trade_tidy)

```

</div><br>

<div style="border: 3px solid #4A9586;">
<span style="color:green">&nbsp;**Summarise the data**</span>
```{r}
trade_summary <- trade_tidy %>% group_by(Period, Type, TradeType) %>% summarise(Avg=mean(Trades)) %>% arrange(desc(Avg))
kable(trade_summary)

ggplot(trade_summary, aes(x=TradeType, y=Avg, fill=TradeType)) + geom_bar(stat = "identity") + facet_grid(Type ~ Period) + labs(x = "Export/Import", y = "Average") + ggtitle("Three Year(2013-2014) Average Span")
```

<span style="color:green">&nbsp;**From examining the plot above, it shows more services are exported than goods. However number of goods exported and imported are much higher than the services.**</span>

</div>

### <span style="color:#4A9586;">1. Calculate the average goods export for each month over the three-year span.</span>
<div style="border: 3px solid #4A9586;">

```{r}
kable(trade_summary %>% filter(Type == "Goods" & TradeType == "Exports"))
```

</div>

### <span style="color:#4A9586;">2. Calculate the average goods import for each month over the three-year span.<span>
<div style="border: 3px solid #4A9586;">
```{r}
kable(trade_summary %>% filter(Type == "Goods" & TradeType == "Imports"))
```

</div>

### <span style="color:#4A9586;">3. Calculate the average services export for each month over the three-year span.</span>
<div style="border: 3px solid #4A9586;">

```{r}
kable(trade_summary %>% filter(Type == "Services" & TradeType == "Exports"))
```

</div>
### <span style="color:#4A9586;">4. Calculate the average services import for each month over the three-year span.</span>

<div style="border: 3px solid #4A9586;">

```{r}
kable(trade_summary %>% filter(Type == "Services" & TradeType == "Imports"))
```

</div> <br><br>