---
title: "607 Assignment 10 - Web APIs"
author: "Puneet Auluck"
date: "November 1, 2015"
output: html_document
---

***


<span style="color:blue"> **Assignment overview:** </span> Download JSON data provided through New York Times Web API portal into R data frame.

<span style="color:blue"> **Data of choice:** </span> New York Times Best Sellers Books


### Step 1: Load required libraries
```{r message=F, warning=F}
library(jsonlite)
library(dplyr)
library(knitr)
```

### Step 2: Create wrapper function getBestSeller
```{r}
#Get books ranked 1st for the type provided in the argument
  getBestSeller <- function(uri, format=".json", list_type, rank=1,api_key){
      json_output    <- fromJSON(paste(uri, format,
                                 "?list-name=", list_type,
                                 "&rank=", rank,
                                 "&api-key=", api_key, 
                                 sep=""))
      results       <- json_output$results
      # book_detail contains list so convert it to data frame and then bind it to results
      details       <- as.data.frame(results$book_detail)
      final_output  <- cbind(results, details) %>% select(list_name, rank,title, author, description)
    
      return(final_output)
}


```

### Step 3: Get most recent Best Sellers

<span style="color:blue"> Store api-key and base url into variables </span>
```{r}
nyt_api_key <- "543adc61e14cf137f5a2121ec35d7d35:4:73310298"
nyt_uri <- "http://api.nytimes.com/svc/books/v3/lists"
```

<span style="color:blue"> Call getBestSeller one at a time </span>
```{r}
# get hardcover-fiction
kable(getBestSeller(uri = nyt_uri,list_type = "hardcover-fiction", api_key = nyt_api_key))

# e-book-fiction
kable(getBestSeller(uri = nyt_uri,list_type = "e-book-fiction", api_key = nyt_api_key))

```

<span style="color:blue"> Or create a vector of items and then call the function </span>
```{r}
list_names <- c("hardcover-fiction","e-book-fiction","trade-fiction-paperback", "hardcover-nonfiction")
best_sellers <- lapply(list_names, function(x) getBestSeller(uri = nyt_uri,list_type = x, api_key = nyt_api_key))
best_sellers_df <- dplyr::rbind_all(best_sellers)
kable(best_sellers_df)

```

***
***