---
title: "R Assigment - Week 2"
author: "Puneet Auluck"
date: "July 9, 2015"
output: html_document
---

###Assignment description
R has built in functions for choose() and factorial(). Your task is to write your own variation of the **choose()** function, that does not use either of these functions.

In probability, "n choose r" means $$ n! / (( n - r)! * r!) $$

Your code should adhere to the style guidelines provided by Hadley Wickham in this week's reading. [http://adv-r.had.co.nz/Style.html]

You need to create your code and some test code, and deliver your code, results, and documentation in a single R Markdown file.

----


### my\_choose() function - uses my\_factorial() function

```{r}

# my_choose function -----------------
# n must be greater than 0 and r should be less than or equal to n
my_choose <- function(n,r){
  
  if (n<0 | r <0){ print("please provide positive integer")}
  if (n < r) { return(0) }
  if (n == r) { return(1) }
  if (n == 0 | r == 0) { return(1) }
  
  if (r %% 1 !=0){
    warning("Second number ", r, " is rounded to: ", round(r))
    r <-round(r)}
  return(my_factorial(n) / (my_factorial(n-r) * my_factorial(r)))
}

```

### my\_factorial() function
```{r}
# my_factorial function --------------
my_factorial <- function(n){
  if( n == 0) { return(1) }
  if(n < 0) { return(0) }
  y <- 1
  for(x in n:1){ y <- y*x }
  return(y)
}

```

###Test Cases (total 4)\: *my_choose()* results are verifed with *choose()* results
**1. Test with valid arguments**
```{r}
c(my_choose(10,5),choose(10,5))
c(my_choose(4,1),choose(4,1))
c(my_choose(7,3),choose(7,3))
```

**2. Test for zero value arguments**
```{r}
c(my_choose(0,0),choose(0,0))
c(my_choose(1,0),choose(1,0))
c(my_choose(0,1),choose(0,1))

```

**3. Test for argument where r > n**
```{r}
c(my_choose(1,17),choose(1,17))
```

**4. Test for arguments that contain fraction values**
```{r}
c(my_choose(7,3.5),choose(7,3.5))

c(my_choose(7.5,3),choose(7.5,3))

c(my_choose(7.5,3.5),choose(7.5,3.5))

```
